\documentclass{article}
\usepackage{graphicx}
\usepackage{enumitem}

\title{Sensor Selection for Ball-on-Beam Project}
\author{Ahmed Ibrahim Almohamed \\ Abdullah Albandar Alhamad}
\date{August 2023}

\begin{document}

\maketitle

\section{Introduction}

The successful realization of the ball-on-beam project necessitates careful sensor selection, with a focus on accuracy and precision. In this document, we explore different sensor options, with a spotlight on Time-of-Flight (ToF) sensors.

\section{Sensor Comparison}

\begin{table}[htbp]
\centering
\begin{tabular}{|p{5cm}|p{6cm}|}
\hline
\textbf{Sensor Type} & \textbf{Description} \\
\hline
Infrared (IR) Distance Sensors & Emit infrared light to measure reflection time, suitable for distance measurement and object presence detection. \\
\hline
Ultrasonic Distance Sensors & Emit sound waves and measure reflection time, facilitating non-contact distance measurement. \\
\hline
Time-of-Flight (ToF) Sensors & Utilize laser/LED light to measure reflection time, enabling highly accurate distance calculations. \\
\hline
Laser Distance Sensors & Employ laser light for precise distance measurements; offer elevated accuracy with potential cost implications. \\
\hline
Reflective Opto Sensors & Detect object presence through light reflection, valuable for relative object positioning. \\
\hline
Capacitive/Inductive Proximity Sensors & Detect object presence without physical contact, suitable for defined range detection. \\
\hline
\end{tabular}
\caption{Comparison of Different Sensor Types}
\label{tab:sensors}
\end{table}
\section{ToF Sensors: Ideal Choice}

Among the sensor types evaluated, ToF sensors stand out as a promising choice for our ball-on-beam project.

\subsubsection{Principle and Advantages}

ToF sensors employ light pulses to measure distances, excelling in accuracy and real-time responsiveness.

\subsubsection{Applications}

The capabilities of ToF sensors make them well-suited for applications such as:

\begin{itemize}
    \item Precise object detection and obstacle avoidance.
    \item Reliable position sensing in confined spaces.
    \item Dynamic environments requiring real-time measurements.
\end{itemize}


\subsection{Conclusion}

After thorough assessment, we've selected ToF sensors as the optimal choice for our ball-on-beam project. These sensors provide accurate, real-time distance measurements, aligning seamlessly with our project's precision requirements.

\subsubsection{Educational Value}

In addition to its technical advantages, the choice of ToF sensors also serves an educational purpose. By opting for ToF sensors, we aim to enhance our understanding of sensor technology and gain practical experience in utilizing such advanced sensing methods.

\end{document}
